### 2016-09-23 review/response_to_reviewer.txt
update review

#+TITLE:  "Heterogeneous Firing Responses Predict Diverse Couplings to Presynaptic Activity in mice Layer V Pyramidal Neurons" \\ _\textsc{ Response to reviewers}_
#+AUTHOR: Y. Zerlaut & A. Destexhe

* Impact of active dendritic mechanisms

Two reviewers raised the concern that active dendritic mechanisms
could qualitatively affect the claim of our study. We address their
concern in the following.

How could active dendritic mechanisms "qualitatively affect the claim
of our study" ? Our framework for single cell computation consists in
(i) evaluating analytically the somatic subthreshold membrane
potential fluctuations at the soma as a function of the presynaptic
quantities and (ii) convert those fluctuations into a spiking
probability thanks to a /firing response/ function determined /in
vitro/ in individual neuron thanks to somatic recordings.

Our framework thus needs to be tested against two phenomena: (I)
whether active dendritic mechanisms qualitatively affect the
relationship between presynaptic quantities and somatic membrane
potential fluctuations and (II) whether dendritic non-linearities
introduce a coupling between presynaptic activity and spike emission
that render the decomposition into the successive steps i & ii
misleading. Those phenomena will naturally appear to some extent and
will produce quantitative deviations with respect to our predictions
(that neglects dendritic non-linearities), the purpose of this study
is only to check for "qualitative deviations".

To investigate the impact of those phenomena, we included active
dendritic mechanisms (as provided by the model of [[citetext:Larkum2009]],
available on ModelDB, [[https://senselab.med.yale.edu/modeldb/showModel.cshtml?model=124043][link]]) into our framework for cellular
computation in the /fluctuation-driven/ regime. We included two of the
principal ingredients of active dendritic mechanisms into our model:
(i) NMDA currents (coupled to the AMPA synapses) and (ii) Ca^{2+}
spikes. We performed a numerical study on the /mean/ morphological
model of the main text (Table 1 and Figure 2,3), we spread those
mechanisms according to the picture provided by the study of
[[citetext:Larkum2009]], all glutamatergic synapses were replaced by the
AMPA+NMDA synapses and Ca^{2+} currents were included in the last two
branches to provide a /relay/ for far dendritic events (depicted in
the mean model drawing of [[fig:0a][Figure 1]]). Phenomenon (I) will be tested
against the inclusion of the NMDA mechanism and phenomenon (II) will
be tested against the inclusion of both NMDA and Ca^{2+} spikes. Those
two situations will be compared to the picture described in the main text.

#+NAME: fig:0a
#+CAPTION: *Including dendritic non-linearities (numerically) within our framework for single cell computation. On top of each panel, we present the recording of the membrane potential in the far distal dendrite (depicted with the upper electrode on the drawing) in the Control and Control+NMDA case. On the bottom, we present the membrane potential recorded at the soma (depicted with the lower electrode on the drawing) in four cases, the Control, the Control+NMDA case, the Control case with a HH model inserted in the soma and the Control+NMDA+Ca^{2+}spikes+HH(soma)*. *(A)* Response to baseline activity. \(\nu_e^p=\nu_e^d\)=0.2Hz, \(\nu_i^p=\nu_i^d\)=1.2Hz and s=0.05. *(B)* Response to baseline activity with the synchronous activation of 20 synapses every 500ms (marked by a triangle), note the impact of the NMDA mechanism under this stimulation type. *(C)* Response to an increase of proximal activity with respect to the baseline level (\(\nu_e^p\)=1.6Hz, \(\nu_i^p\)=10Hz). *(D)* Response to an increase in distally targetting presynaptic activity. *(E)* Response to an increase in unbalanced activity. *(F)* Response to an increase in 
[[./figs/traces.pdf]]

In [[fig:0a][Figure 1]], we represent the output of the numerical simulations of
the resulting model. We show on [[fig:0a][Figure 1]]A,C-F the response to the five
protocols discussed within the manuscript. Additionally, on [[fig:0a][Figure 1]]B,
we added another type stimulus

[[cite:Stuart2015]]

show the their impact in the
fluctuation-driven regime, see [[fig:0b][Figure 2]].

#+NAME: fig:0b
#+CAPTION: *Effect of dendritic non-linearities in the subthreshold fluctuations and firing response in the fluctuation-dirven regime.* *(A)* 
[[./figs/active_sum_up.pdf]]

* Reviewer #1

\it Unfortunately, it is very difficult to grasp the essence and even more
the details of the study. \rm

1) \it Main statement; presentation. It starts with the title that claims
   that claims that “Heterogeneous Firing Responses Predict Diverse
   Couplings to Presynaptic Activity in mice Layer V Pyramidal
   Neurons”. The title refers to the phenomenology in Figure 5 showing
   that the output frequency goes up or down for different cells as a
   function of “increasing presynaptic quantity”, as the x-axis in
   that figure is label. But not clear how strong the increase is at
   the corresponding position, nor what of the presynaptic quantity
   is. For us (the writer and two other group members) it was neither
   possible to figure it out from the caption nor the text. \rm

    We agree that this way to represent data might not be
   straightforward. We improved the legend of Figure 4 to facilitate
   the comprehension. In particular, we have added an example to
   explicit the presence of "comodulations" of variables and hence the
   difficulty to represent the data as a function of one given
   variable. We believe that this way to present data is still the
   best one to emulate the comodulations that characterize neocortical
   dynamics (e.g. inhibition follows excitation) and to be able to
   represent the response to various forms of network activity
   (i.e. various comodulations) on individual cells.  The new legend
   reads:

   _The abscissa “increasing synaptic quantity” corresponds to the comodulations of the model variables shown in Figure 4A (same color code). As an example, the response to a “proximal activity increase” (blue curve) corresponds to a linear increase of and while keeping , and to their baseline level. See Figure 4A for the comodulations of the other types of increasing activity._

2) \it Main statement unclear. Unfortunately, it does not become clear
   what one has to understand of the “Diverse Couplings to the
   Presynaptic Activity” as expressed in the title. In the abstract it
   is stated: “We indeed found such a heterogeneous response, but
   trivially, the only significant contribution to the various firing
   response was the different levels of cellular excitability.” We
   guess that the “trivially” refers to the fact that only the
   cellular excitability has been modulated in the model. But then we
   don’t understand “A notable exception appeared for proximal
   activity: increasing proximal activity could either promote firing
   response in some cells or suppress it in some other cells whatever
   their individual excitability.” We conclude that there is still
   more than the individual excitability that determines the firing
   behavior, but this seems to contradict the previous statement. \rm

   We agree with the reviewer that the formulation "only [...] except
   [...]" might be misleading, we adapted the abstract to remove this
   source of confusion. Now it reads: "but trivially, the _main_
   significant contribution to the various firing response was the
   different levels of cellular excitability [...]".

#+NAME: fig:1
#+CAPTION: *Changing the ratio of the proximal/distal separation.* *(A)* Configuration of the main text, the proximal part occupies the first 87.5% of the dendritic morphology. *(B)* Now the proximal part occupies only the first 12.5% of the dendritic morphology. This importantly reduces the number of available synapses (the morphology has only one branch in the first 12.5% part). Provided this loss of synapses is compensated by an increase of activity, there is _no qualitative change between the 12.5% and the 87.5% setting for the proximal activity: non-monotonic \(\sigma_V\) variations, strong decrease in \(\tau_V / \tau_m^0 \)_. On the other hand, this change would have important qualitative consequences for the "distal activity" regime, if the last 87.5% of the morphology are "distal", many synapses are electrotonically close (what is rather irrealistic for distal synapses), they will thus stronly interact through conductance shunting and thus produce fluctuations close to the "proximal activity" setting of the main text.
[[./figs/varying_prox_dist.png]]

3) \it Main statement; missing analysis. It is suggested that the
   heterogeneity would arise from sodium channels, but the paper is
   missing a joint presentation of experimental and modeling data
   showing some fit. Neither could we find a mathematical argument why
   the different phenomenological behavior shown in the cited Fig. 5
   should arise. It would be helpful to underscore the main statement
   by a simplified mathematical formulation in the main text. For
   instance, a simplified formula that gives the output firing rate as
   a function of the input variables would be helpful, including the
   dependencies on the sodium channel density. This sodium channel
   density is claimed to make the difference between the cell types
   according to the abstract: “various levels of density of sodium
   channels and sodium inactivation will control the response of
   individual neurons to proximally targeting presynaptic activity.”
   Again, we could not find a state-of-the-art explanation why this
   statement is true. \rm

    This point is precisely the focus of our previous communication
   [[cite:Zerlaut2016]]. Unfortunately there is, to the best of our knowledge, no
   closed-form solution that could give the firing rate response as a
   response to fluctuating input in a model including sodium
   inactivation and a sdium activation curve (hence the
   phenomenological approach formulated in ou previous
   communication). Nonetheless to give some intuition to the reader,
   we have added the following statement (line 316):
   _In our previous communication, theoretical modeling suggested that a high sensitivity to the speed of the fluctuations was enable by a high level of sodium inactivation (as only fast fluctuations allow to deinactivate sodium channels) and a high density of sodium channels (as it  produces a sharp spike initiation that enables to extract fast fluctuating input)._
   For further details, we believe the reader should refer to our
   previous publication.

4) \it No dendritic nonlinearities. One main characteristics of L5
   pyramidal neurons is the calcium spike in the apical tree, and more
   generally the various types of dendritic spikes. Non of them is
   considered in the model, and one conjectures that the various
   response properties in the experimental data arise from these
   various dendritic nonlinearities. Describing the firing properties
   of pyramidal neurons as a function of proximal and distal activity
   should take account of these dendritic nonlinearities. Otherwise,
   in fact, the statements cannot be compelling. \rm

   We adressed this concern in details in Part A.

5) \it Data extraction. It is not clear how the 4 cells have been selected
   out of the 27 cells. The 4 cells represent 4 cell types. How are
   these cell types specified? The paper does not give an answer on
   that. \rm

   The 4 cells represent 4 different cells out of the n=30 cells
   recorded in our previous study [[cite:Zerlaut2016]]. The criteria to
   choose those 4 cells was that they should somehow represent
   different behaviors while still being representative. See in Figure
   6B in [[citetext:Zerlaut2016]] how those 4 cells cover relatively well
   the domain of measured characteristics for the firing response,
   this is also the case in the present study, where they cover
   relatively well the domain of the predicted coupling to the various
   types of activities.

6) \it Figure 2 has two panels A & B, but the caption describes panels
   A–E. \rm

   Indeed, we apologize for this problem in the manuscript
   preparation. Figure 2 and Figure 3 have been mixed up...

7) \it Apart from that, the fit of the numerical simulations and the
   analytical estimate in B for sigma_V is biased. No argument is
   given why this bias is there. \rm

   From line 165 to line 182, we discuss the validity of our
   analytical treatment. We explicit the two hypothesis that underlie
   our estimation and detail their expected effect. We conclude by
   discussing the bias for \(\sigma)V\) from line 180 to 182: "Some
   discrepancies between the approximation and the simulations
   appeared, in particular one can see a 1mV shift in the standard
   deviation of the fluctuation (meaning that single events are
   underestimated in the analytical treatment, so that hypothesis 2 is
   the most problematic one)." (where hypothesis 2 has been explicited
   before).

8) \it The central claim is that distal and proximal inputs differently
   affects the output firing rates. Beside the fact that this
   statement by itself is trivial [...] \rm

   We also show that "distal and proximal inputs differently
   affects the output firing rates", but this is not our main
   statement and we agree with the reviewer that this could seen as
   "trivial". Our main statement claims that those stimulations affect
   individual neurons in diverse manners (in particular for the
   proximal stimulation: increasing their response for some neurons,
   decreasing it for some other neurons). We believe that this
   diversity of responses across neurons (to the same input) is not a
   so trivial phenomena.

9) \it [...] we wondered why the closer 87.5% of the dendritic branch is
   declared as “proximal” and the furthest 12.5% as the “distal”
   part. Is the result dependent on this choice, or could one have
   also chosen the closer 12.5% as “proximal” only? \rm

   Indeed, the _arbitrary ratio_ between proximal and distal input
   might need a more detailed discussion. Following the reviewer's
   suggestion, we reproduced the analysis with the convention
   "proximal as the closest 12.5% only". We show on [[fig:1][Figure 3]] that the
   results are not affected by this change, the qualitative properties
   of the fluctuations remain identical: strong decrease of the
   fluctuations time constant and non-monotonic shape of the standard
   deviation of the fluctuations. The key property leading to the
   properties of the "proximal activity" being that they correspond to
   synapses that are close to each others and to the soma (to produce
   a strong /shunting/ effect), this phenomena is even stronger under
   this setting. The reason for choosing a relatively large fraction
   for proximal synapses is to obtain a large number of
   synapses. Indeed, along the morphology, the electrotonic distance
   raises but the surface as well (because of the branches number, see
   drawing in Fig 1 in main text) so that the number of synapses
   strongly increases. Summing up, the parameter \(f_{prox}\)=87.5% is
   important to keep the distal part far enough electrotonically, for
   the proximal part there is no much difference between
   \(f_{prox}\)=87.5% and \(f_{prox}\)=12.5% (they all contain the
   very proximal synapses), and keeping \(f_{prox}\)=87.5% avoid to
   introduce an additional rescaling of synaptic frequences because of
   the reduced number of synapses (the procedure shown in [[fig:1][Figure 3]]).

* Reviewer #2

 This manuscript shows a nice mixture of experiment and theory. The
 latter is based upon sensible use of the cable modelling framework of
 Rall combined with stochastic models of synaptic input (admittedly of
 a particular type but definitely fit for purpose).

\bf we thank the reviewer for his \rm

1) Unfortunately some of the work (for the firing response function)
   is buried in supplementary material - and this has not been made
   available to me (S1).

   \bf Sorry to hear this. For us, the link at the end of the
   submission pdf (page 32/32) was functional and actually gave the
   supplementary material. \rm

2) It would be especially informative to hear more about the author's
   perspective on how synchrony controls rate. Nonetheless given the
   quality and novelty of the main manuscript I suggest to put this
   issue aside.

   \bf \rm

** Major concern

My only major concern with the paper is on the emphasis of a
theoretical framework without citing other relevant work in the
area. Essentially I am surprised not to see the mention of the
technique(s) for constructing the Green's function of an arbitrary
branched passive dendritic tree. The paper focuses on Rall equivalent
cylinders for simplicity, yet the methodology presented would seem
easily extended using the work of Abbott [[cite:Abbott1991,Cao1993]]. The
extension of the above to include a distinguished soma model can be
found in [[citetext:Coombes2007]] and Paul Bressloff has done a lot of
work on dendrites that appears relevant, some of which is summarised
in [[citetext:Bressloff1997]]. The following may also be of interest
(though not necessarily for this work) - where simple active soma
models are coupled to passive dendrites
[[cite:Bressloff1995,Svensson2009,Schwemmer2012]].

\bfseries

Indeed, our

We thank the reviewer

\textnormal

** Minor points

Spurious "s" in the sec "A simplified morphological model" - sWe arbitrarily ...
Where is excitability < V_thre^eff >_D defined?

\bf We apologize for the problems with the manuscript preparations and we thank the reviewer for pointing out those mistakes. \rm

* Reviewer #3

\it The key claim of this work is that recently identified L5 PCs firing
rate properties indicate how these cells couple to presynaptic
activity patterns (proximal, distal,...). The claim is based on i)
recently identified firing response functions (previous work of the
authors), ii) a theoretical model for passive dendritic integration
that is an extension of the single compartment model of
[[citetext:Kuhn2004]] to a passive tree with cable theory, and iii) an
identification of the passive tree of L5 PCs by fitting the input
impedance at the soma to experimental data.

The topic is of importance and the theoretical part is novel, very
appealing in itself and rigorous, but inadequate to support the key
claim of the paper (see paper title).

This is because ... \rm

** CONS

 - *A*: (part ii) For L5 PCs it is highly questionable if dendritic
   integration can be understood without (or separated from) active
   dendritic properties given the known experimental facts and huge
   literature, e.g. about the compartmentalization of dendritic
   function, dendritic spikes, and changed response properties due to
   ion channel blockers.

   In my opinion, the model for the L5 PCs based on a passive tree
   will almost surely fail to describe the real dendritic integration
   properties of the cells in the fluctuation driven regime. In any
   case, see the first point in additional work below.

- *B*: (part iii) Even the description of the passive dendritic
  integration properties of the cell seems to be questionable (maybe
  because of the constraints to symmetrical trees). That this doesn't
  work out properly seems to be reflected in the morphologies obtained
  in the paper: Is the required variation of the tree length from
  about 400 um to 900 um (current Fig 3C and Fig. S1) for PCs
  plausible? Please provide references. L5 PCs with their long apical
  dendrites etc. are the worst case scenario for the symmetry
  constraints for the tree.

** PROS

The theoretical work is still significant for other cell types that
work in a more linear regime, e.g. L2/3 PCs, and have morphologies
that match the theoretical constraints..

** Additional work required:

- The authors should provide evidence that active dendritic
  integration doesn't change the key claim, or, apply the model to a
  more appropriate cell type.

- Because the morphology of the tree of the model is very constraint
  and only a very rough estimate its necessary to assess the influence
  of this approximation on dendritic integration: A systematic
  analysis of the impact of the morphology of the tree (7 variables)
  on the response properties of interest at the soma is missing (This
  allows also a generalization of the results to other cell types. Now
  only Fig S2 gives a hint about variability).

- There's no reference to any other model than [[citetext:Kuhn2004]], on
  which the approach is based. Relate the model to recent work on the
  effects of passive dendritic tree properties of neuron on firing
  dynamics (e.g.  [[citetext:Saparov2015]] or other model reduction
  techniques (a huge field, e.g. [[citetext:Hedrick2013]]).

The paper is well organized, easy to read, simple to follow and
clearly written. However, there are too many typos and word
repetitions to list them all ("sWe" p.3, "show deviations from the
deviations" p.4, " vary the size of the size" p.4, "does not goes",
p.6, "the firing increase show", p.7 ...) and the Figure numbers were
incorrect (Fig.22b, p.4) or mixed up (Fig. 2 and 3).

Minor:

- Fig. 3: The color code seems to be inconsistent: Panel A shows cells
  with Rm above 600 MOhm, but the bottom panel in Fig 3C shows that
  such cells are not present.

- Fig 3C: The lines in the three top panels could be drawn with the
  color code (color changes within a line) as used in panel A and B.

- Fig 3AB: Please comment on the inconsistency between data and model
  for > 100 Hz. There seems to be a flip in vertical order.

Section: "An analytical approximation ...":

- The claim "the analytical estimate could therefore be seen as an accurate estimate, module a shift in the synchrony (~0.2 increase in synchrony corrects for the ~1 mV shift in sigma)" should verified:
1) Results for synchrony = 0.2 should be shown in the right most column of panels in Fig 2B, and
2) results for synchrony = 0.2 should be shown for the other columns of panels in Fig 2B, where the other parameters are varied.

Section: "Biophysical origin ..."

- The characteristics of the firing response function are not properly introduced. Please add this.

- It can be hardly claimed that Fig. 6A is approximately a log normal distribution.

Methods section:

- Not all variables are properly defined, e.g. Qe, Qi, lp.

- "it constitutes an extension of the approach proposed in Kuhn et al. \[19\]": Please, state what extension.

- "Fitting passive properties": Are the solutions unique or are there many different trees with similar fit performance that result in different couplings to presynaptic activity?

* References
\small

#+BEGIN_LATEX
\begin{filecontents}{biblio.bib}

@article{Stuart2015,
author = {Stuart, Greg J and Spruston, Nelson},
doi = {10.1038/nn.4157},
file = {:Users/yzerlaut/Documents/Mendeley Desktop/Stuart, Spruston - 2015 - Dendritic integration 60 years of progress.pdf:pdf},
issn = {1097-6256},
journal = {Nature Neuroscience},
mendeley-groups = {Neuroscience},
number = {12},
pages = {1713--1721},
pmid = {26605882},
title = {{Dendritic integration: 60 years of progress}},
url = {http://www.nature.com/doifinder/10.1038/nn.4157},
volume = {18},
year = {2015}
}

@article{Larkum2009,
author = {Larkum, M E and Nevian, T and Sandler, M and Polsky, A and Schiller, J},
doi = {10.1126/science.1171958},
issn = {1095-9203 (Electronic) 0036-8075 (Linking)},
journal = {Science},
number = {5941},
pages = {756--760},
title = {{Synaptic integration in tuft dendrites of layer 5 pyramidal neurons: a new unifying principle}},
type = {Journal Article},
url = {http://www.ncbi.nlm.nih.gov/pubmed/19661433 http://science.sciencemag.org/content/sci/325/5941/756.full.pdf},
volume = {325},
year = {2009}
}

@article{Hedrick2013,
author = {Hedrick, K R and Cox, S J},
doi = {10.1007/s10827-012-0403-y},
issn = {1573-6873 (Electronic) 0929-5313 (Linking)},
journal = {J Comput Neurosci},
number = {1},
pages = {1--26},
title = {{Structure-preserving model reduction of passive and quasi-active neurons}},
type = {Journal Article},
url = {https://www.ncbi.nlm.nih.gov/pubmed/22714391},
volume = {34},
year = {2013}
}

@article{Saparov2015,
  title={Effects of passive dendritic tree properties on the firing dynamics of a leaky-integrate-and-fire neuron},
  author={Saparov, Abulhair and Schwemmer, Michael A},
  journal={Mathematical biosciences},
  volume={269},
  pages={61--75},
  year={2015},
  publisher={Elsevier}
}

@article{Schwemmer2012,
  title={Bistability in a leaky integrate-and-fire neuron with a passive dendrite},
  author={Schwemmer, Michael A and Lewis, Timothy J},
  journal={SIAM Journal on Applied Dynamical Systems},
  volume={11},
  number={1},
  pages={507--539},
  year={2012},
  publisher={SIAM}
}

@article{Bressloff1995,
  title={Dynamics of a compartmental model integrate-and-fire neuron with somatic potential reset},
  author={Bressloff, Paul C},
  journal={Physica D: Nonlinear Phenomena},
  volume={80},
  number={4},
  pages={399--412},
  year={1995},
  publisher={Elsevier}
}

@article{Svensson2009,
author = {Svensson, C.-M. and Coombes, S},
doi = {10.1142/S0218127409024347},
journal = {International Journal of Bifurcation and Chaos},
pages = {2597--2607},
title = {{Mode Locking in a Spatially Extended Neuron Model: Active Soma and Compartmental Tree}},
type = {Journal Article},
volume = {19},
year = {2009}
}

@article{Bressloff1997,
author = {Bressloff, P. C. and Coombes, S.},
doi = {10.1142/S0217979297001209},
file = {:Users/yzerlaut/Documents/Mendeley Desktop/Bressloff, Coombes - 1997 - Physics of the Extended Neuron.pdf:pdf},
issn = {0217-9792},
journal = {International Journal of Modern Physics B},
mendeley-groups = {Neuroscience},
month = {aug},
number = {20},
pages = {2343--2392},
title = {{Physics of the Extended Neuron}},
url = {http://www.worldscientific.com/doi/abs/10.1142/S0217979297001209},
volume = {11},
year = {1997}
}

@article{Coombes2007,
author = {Coombes, S and Timofeeva, Y and Svensson, C-M and Lord, G J and Josi{\'{c}}, K and Cox, S J and Colbert, C M},
doi = {10.1007/s00422-007-0161-5},
file = {:Users/yzerlaut/Documents/Mendeley Desktop/Coombes et al. - 2007 - Branching dendrites with resonant membrane a sum-over-trips approach.pdf:pdf},
issn = {0340-1200},
journal = {Biological cybernetics},
keywords = {Action Potentials,Action Potentials: physiology,Algorithms,Animals,Cell Membrane,Cell Membrane: physiology,Cell Shape,Cell Shape: physiology,Computer Simulation,Dendrites,Dendrites: physiology,Hippocampus,Hippocampus: physiology,Membrane Potentials,Membrane Potentials: physiology,Models, Neurological,Rats,Synaptic Transmission,Synaptic Transmission: physiology},
mendeley-groups = {Neuroscience},
month = {aug},
number = {2},
pages = {137--49},
pmid = {17534649},
title = {{Branching dendrites with resonant membrane: a "sum-over-trips" approach.}},
url = {http://www.ncbi.nlm.nih.gov/pubmed/17534649},
volume = {97},
year = {2007}
}

@article{Cao1993,
author = {Cao, B J and Abbott, L F},
doi = {10.1016/S0006-3495(93)81370-5},
issn = {0006-3495 (Print) 0006-3495 (Linking)},
journal = {Biophys J},
number = {2},
pages = {303--313},
title = {{A new computational method for cable theory problems}},
type = {Journal Article},
url = {https://www.ncbi.nlm.nih.gov/pubmed/8457662},
volume = {64},
year = {1993}
}

@article{Abbott1991,
author = {Abbott, L F and Farhi, E and Gutmann, S},
issn = {0340-1200 (Print) 0340-1200 (Linking)},
journal = {Biol Cybern},
number = {1},
pages = {49--60},
title = {{The path integral for dendritic trees}},
type = {Journal Article},
url = {https://www.ncbi.nlm.nih.gov/pubmed/1768712},
volume = {66},
year = {1991}
}

@article{Zerlaut2016,
author = {Zerlaut, Y and Telenczuk, B and Deleuze, C and Bal, T and Ouanounou, G and Destexhe, A},
doi = {10.1113/JP272317},
issn = {1469-7793 (Electronic) 0022-3751 (Linking)},
journal = {J Physiol},
number = {13},
pages = {3791--3808},
title = {{Heterogeneous firing rate response of mouse layer V pyramidal neurons in the fluctuation-driven regime}},
type = {Journal Article},
url = {http://www.ncbi.nlm.nih.gov/pubmed/27146816},
volume = {594},
year = {2016}
}

@article{Kuhn2004,
abstract = {During sensory stimulation, visual cortical neurons undergo massive synaptic bombardment. This increases their input conductance, and action potentials mainly result from membrane potential fluctuations. To understand the response properties of neurons operating in this regime, we studied a model neuron with synaptic inputs represented by transient membrane conductance changes. We show that with a simultaneous increase of excitation and inhibition, the firing rate first increases, reaches a maximum, and then decreases at higher input rates. Comodulation of excitation and inhibition, therefore, does not provide a straightforward way of controlling the neuronal firing rate, in contrast to coding mechanisms postulated previously. The synaptically induced conductance increase plays a key role in this effect: it decreases firing rate by shunting membrane potential fluctuations, and increases it by reducing the membrane time constant, allowing for faster membrane potential transients. These findings do not depend on details of the model and, hence, are relevant to cells of other cortical areas as well.},
author = {Kuhn, Alexandre and Aertsen, Ad and Rotter, Stefan},
doi = {10.1523/JNEUROSCI.3349-03.2004},
file = {:Users/yzerlaut/Documents/Mendeley Desktop//Kuhn, Aertsen, Rotter - 2004 - Neuronal integration of synaptic input in the fluctuation-driven regime.pdf:pdf},
issn = {1529-2401},
journal = {The Journal of neuroscience : the official journal of the Society for Neuroscience},
keywords = {Action Potentials,Action Potentials: physiology,Animals,Computer Simulation,Excitatory Postsynaptic Potentials,Excitatory Postsynaptic Potentials: physiology,Humans,Membrane Potentials,Membrane Potentials: physiology,Models,Neural Inhibition,Neural Inhibition: physiology,Neurological,Neurons,Neurons: physiology,Synapses,Synapses: physiology,Synaptic Transmission,Synaptic Transmission: physiology,Visual Cortex,Visual Cortex: physiology},
mendeley-groups = {Neuroscience},
month = {mar},
number = {10},
pages = {2345--56},
pmid = {15014109},
publisher = {Soc Neuroscience},
title = {{Neuronal integration of synaptic input in the fluctuation-driven regime.}},
url = {http://www.ncbi.nlm.nih.gov/pubmed/15014109},
volume = {24},
year = {2004}
}
\end{filecontents}
#+END_LATEX

\bibliography{biblio}

* Options :noexport:
# #+OPTIONS: toc:nil num:1
#+OPTIONS: num:1
#+LaTeX_CLASS_OPTIONS: [8pt, a4paper, colorlinks]
#+LaTeX_HEADER: \hypersetup{allcolors = blue} % to have all the hyperlinks in 1 color
#+LaTeX_HEADER: \usepackage{amssymb,mathenv,array}
#+LaTeX_HEADER: \usepackage{microtype} % Slightly tweak font spacing for aesthetics
#+LaTeX_HEADER: \usepackage{geometry}
#+LaTeX_HEADER: \geometry{a4paper,total={210mm,297mm}, left=25mm, right=25mm, top=20mm, bottom=20mm}
#+LaTeX_HEADER: \usepackage{filecontents}
#+LaTeX_HEADER: \renewcommand{\refname}{\vspace{-.8cm}}
#+LaTeX_HEADER: \usepackage{biology_citations}
#+LaTeX_HEADER: \bibliographystyle{biology_citations}
#+LaTeX_HEADER: \renewcommand{\thesection}{\Alph{section}}