### 2016-09-27 review/response_to_reviewer.txt
ädding new final check sim theory

#+TITLE:  "Heterogeneous Firing Responses Predict Diverse Couplings to Presynaptic Activity in mice Layer V Pyramidal Neurons" \\ _\textsc{ Response to reviewers}_
#+AUTHOR: Y. Zerlaut & A. Destexhe

* Impact of active dendritic mechanisms

Two reviewers raised the concern that active dendritic mechanisms
(reviewed in [[citetext:Stuart2015]]) could qualitatively affect the claim
of our study. We address their concern in the following.

How could active dendritic mechanisms "qualitatively affect the claim
of our study" ? Our framework for single cell computation consists in
(*i*) evaluating analytically the somatic subthreshold membrane
potential fluctuations at the soma as a function of the presynaptic
quantities and (*ii*) convert those fluctuations into a spiking
probability thanks to a /firing response/ function determined /in
vitro/ in individual neuron thanks to somatic recordings.

Our framework thus needs to be tested against two phenomena: (*I*)
whether active dendritic mechanisms qualitatively affect the
relationship between presynaptic quantities and somatic membrane
potential fluctuations, i.e. checking the validity of step *i*, and
(*II*) whether dendritic non-linearities introduce a coupling between
presynaptic activity and spike emission that render the decomposition
into the successive steps *i* & *ii* misleading. Those phenomena will
naturally appear to some extent and will produce quantitative
deviations with respect to our predictions (that do not include
dendritic non-linearities, e.g. trivially, all cells should be more
excitable). As the main message of the paper does not rely on the
absolute values of the input-output functions but on the rather on the
relative behaviors from cell to cell that emerge from differences
across individual firing response functions, the purpose of this study
is only to investigate putative "qualitative deviations".

To investigate the impact of those phenomena, we included active
dendritic mechanisms (as provided by the model of [[citetext:Larkum2009]],
available on ModelDB on the following [[https://senselab.med.yale.edu/modeldb/showModel.cshtml?model=124043][link]]) into our framework for
cellular computation in the /fluctuation-driven/ regime. We included
two of the principal ingredients of non-linear dendritic mechanisms
into our model: NMDA currents (coupled to the AMPA synapse) and
Ca^{2+} currents. We performed a numerical study on the /mean/
morphological model of the main text (Table 1 and Figure 2,3), we
spread those mechanisms according to the picture provided by the study
of [[citetext:Larkum2009]], all glutamatergic synapses were replaced by
the AMPA+NMDA synapses and Ca^{2+} currents were included in the last
two branches to provide a /relay/ for far dendritic events (depicted
on the drawing of the mean model, see [[fig:0a][Figure 1]]). Phenomenon (*I*) will
be tested against the inclusion of the NMDA mechanism, we will examine
the properties of the membrane potential fluctuations at the soma,
with and without this mechanism. Phenomenon (*II*) will be tested
against the inclusion of both NMDA and Ca^{2+} currents: we will first
include numerically a spike emission model at the soma (Hodgkin-Huxley
like, Na^{+} and K^{+} currents) in the morphological model, then the
model will be able to produce the full input-output function (from
presynaptic quantities to spiking probability), we will then
investigate how NMDA and Ca^{2+} currents impacts the dependency on
the presynaptic variables for the full input-output function.

#+NAME: fig:0a
#+CAPTION: *Including dendritic non-linearities (numerically) within our framework for single cell computation. On top of each panel, we present the recording of the membrane potential in the far distal dendrite (depicted with the upper electrode on the drawing) in the Control and Control+NMDA case. On the bottom, we present the membrane potential recorded at the soma (depicted with the lower electrode on the drawing) in four cases, the Control, the Control+NMDA case, the Control case with a HH model inserted in the soma and the Control+NMDA+Ca^{2+}spikes+HH(soma)*. *(A)* Response to baseline activity. \(\nu_e^p=\nu_e^d\)=0.2Hz, \(\nu_i^p=\nu_i^d\)=1.2Hz and s=0.05. *(B)* Response to baseline activity with the synchronous activation of 20 synapses every 500ms (marked by a triangle), note the impact of the NMDA mechanism under this stimulation type. *(C)* Response to an increase of proximal activity with respect to the baseline level (\(\nu_e^p\)=1.6Hz, \(\nu_i^p\)=10Hz). *(D)* Response to an increase in distally targetting presynaptic activity (\(\nu_e^d\)=0.6Hz, \(\nu_i^d\)=3Hz). *(E)* Response to an increase in unbalanced activity (\(\nu_e^p\)=0.4Hz, \(\nu_i^p\)=2Hz,\(\nu_e^p\)=0.5Hz, \(\nu_i^p\)=1.7Hz). *(F)* Response to an increase in presynaptic synchrony (s=0.4).
[[./figs/traces.pdf]]

#+NAME: fig:0b
#+CAPTION: *Effect of dendritic non-linearities in the subthreshold fluctuations and firing response in the fluctuation-dirven regime.* *(A)* 
[[./figs/active_sum_up.pdf]]

In [[fig:0a][Figure 1]], we represent the output of the numerical simulations of
the resulting models. We represent the subthreshold responses on top
plots, as well as the responses in presence of a spiking mechanism
(the HH model at the soma) on bottom plots. We show on [[fig:0a][Figure 1]]A,C-F
the response to the five protocols discussed within the
manuscript. Additionally, on [[fig:0a][Figure 1]]B, we added another type of
stimulus: the synchronous activation of 20 synapses on one distal
dendrite. The response to this last stimulus clearly exhibits the
presence of strong NMDA currents resembling those observed /in vivo/
[[cite:Larkum2009]]. We observe that, for the /baseline activity/, for the
/strong proximal activity/ and for the /strong distal activity/, there
does not seems to be a strong impact of the NMDA mechanism on the
subthreshold fluctuations (both at the soma and in the distal
dendrites): it exerts a slight depolarization but does not
qualitatively impacts the fluctuations. Neither the triggering of
action potentials seems to be affected (spikes are identically evoked
with and without NMDA and Ca^{2+} channels in presence of a somatic HH
model), see [[fig:0a][Figure 1]]A,C,D. On the other hand, for a high level of
unbalanced activity (where excitation globally overcomes inhibition),
we observe a strong depolarization induced by the NMDA mechanism, we
also observe that, consequently, the spiking probability seems
significantly increased, see [[fig:0a][Figure 1]]E. Also for a high level of
synchrony the fluctuations and spikes seems to be affected, we observe
the presence of small NMDA spikes and additional spikes are elicited
in the presence of NMDA and Ca^{2+} currents, see [[fig:0a][Figure 1]]E.

We then analyze with more details how those active mechanisms impacts
the dependency on the variables of the model. We thus reproduced the
analysis of Figure 3 in the main text. On [[fig:0b][Figure 2]]A, we report the
properties of the fluctuations as a function of the presynaptic
variables, with and without NMDA channels. We observe that NMDA
channels exert an additional small depolarization (increased
\(\mu_V\)), increase the standard deviation of the fluctuations
because they increase the amplitude of depolarizing events (increased
\(\sigma_V\)) and they slow down the fluctuations because they slowly
increase the low frequency content of post-synaptic depolarization
(increased \(\tau_V\)). We also observe, that, despite this small
shift in all quantities, the qualitative trend of all curves remain
identical. Only for high excitation a qualitative difference can be
observe (see bottom left plot of [[fig:0b][Figure 2]]A): increasing excitation
makes V_m fluctuations faster in the passive case whereas increasing
excitation makes fluctuations slower in the presence of NMDA currents
(decreasing vs increasing \(\tau_V\) curve respectively). We conclude
that phenomenon *I* has a very weak impact provided excitation does
not overcome inhibition and provided that synchrony remains in a
relatively low range (the range of our study: $s \leq 0.4\)).

On [[fig:0b][Figure 2]]B, we investigate how NMDA and Ca^{2+} currents shape the
dependency of the spiking probability on the presynaptic
variables. Trivially, we observe an increased spiking probability
because cells are more excitable (NMDA and Ca^{2+} currents are
depolarizing currents). Notably, when excitation does not balance
inhibition (\(\nu_e^p = 0.4\)Hz or \(\nu_i^p = 0\)Hz), the spiking
probability is greatly enhanced (almost doubled). For the other
variables, those effects are much more moderate. Here again, we
conclude that phenomenon *II* has a weak impact provided excitation
does not overcome inhibition and provided that synchrony remains in a
relatively low range (the range of our study: $s \leq 0.4\)).

As can be seen in [[fig:0a][Figure 1]]B, those results nonetheless apply only to
neocortical processing when effects are mediated by /weakly
synchronous and relatively balanced/ population dynamics, when single
cell response is driven by synchronous activation of multiple
synapses, dendritic nonlinearities have a crucial impact (consistent
with experimental evidences and the current view on dendritic
processing, reviewed in [[citetext:Stuart2015]]). We thus modified the
main text to briefly mention the results of the present analysis and
our conclusions. The main text now reads:

\bf The proposed theoretical framework for single-cell computation
nonetheless suffers from several weaknesses. [...]. Most importantly,
there is no active mechanisms in dendrites [[cite:Hausser2005]], it is
therefore a question how much those mechanisms could affect the
picture provided in our study. Preliminary numerical analysis
performed in presence of NMDA and Ca2+ currents [[cite:Larkum2009]],
showed that, provided excitation balances inhibition (a situation
where NMDA channels keeps a relatively low level of stationary
activation) and provided synchrony do not reach a too high level (so
that excitatory events have a weak probability to evoke NMDA and
Ca^{2+} spikes), the qualitative behavior of the cellular input-output
function remains identical. Our preliminary analysis thus suggest the
hypothesis that active dendritic mechanisms may not be crucial for
effects mediated by slow (T\(\geq\)20-30ms) variations of weakly synchronous
population dynamics while being critical for operations performed on
synchronous synaptic activations (T\(\leq\)5-10ms), consistent with the
current view on dendritic processing [[cite:Stuart2015]]. Future work
should address this hypothesis in details. \rm

Because we believe that those results constitute only preliminary data
and that this problem should be the focus of a much more detailed
analysis, we do not wish to include it in the main text or
supplementary material. If the reviewer think that this is important
for our study, we can merge [[fig:0a][Figure 1]] and [[fig:0b][Figure 2]] and produce a
supplementary figure whose legend will be adapted from this section
and the current legends.

* Reviewer #1

\it Unfortunately, it is very difficult to grasp the essence and even more
the details of the study. \rm

1) \it Main statement; presentation. It starts with the title that claims
   that claims that “Heterogeneous Firing Responses Predict Diverse
   Couplings to Presynaptic Activity in mice Layer V Pyramidal
   Neurons”. The title refers to the phenomenology in Figure 5 showing
   that the output frequency goes up or down for different cells as a
   function of “increasing presynaptic quantity”, as the x-axis in
   that figure is label. But not clear how strong the increase is at
   the corresponding position, nor what of the presynaptic quantity
   is. For us (the writer and two other group members) it was neither
   possible to figure it out from the caption nor the text. \rm

    We agree that this way to represent data might not be
   straightforward. We improved the legend of Figure 4 to facilitate
   the comprehension. In particular, we have added an example to
   explicit the presence of "comodulations" of variables and hence the
   difficulty to represent the data as a function of one given
   variable. We believe that this way to present data is still the
   best one to emulate the comodulations that characterize neocortical
   dynamics (e.g. inhibition follows excitation) and to be able to
   represent the response to various forms of network activity
   (i.e. various comodulations) on individual cells.  The new legend
   reads:

   \bf The abscissa “increasing synaptic quantity” corresponds to the
   comodulations of the model variables shown in Figure 4A (same color
   code). As an example, the response to a “proximal activity
   increase” (blue curve) corresponds to a linear increase of
   \(\nu_e^{prox}\) and \(\nu_i^{prox}\) while keeping
   \(\nu_e^{dist}\), \(\nu_i^{dist}\) and \(s\) to their baseline
   level. See Figure 4A for the comodulations of the other types of
   increasing activity. \rm

2) \it Main statement unclear. Unfortunately, it does not become clear
   what one has to understand of the “Diverse Couplings to the
   Presynaptic Activity” as expressed in the title. In the abstract it
   is stated: “We indeed found such a heterogeneous response, but
   trivially, the only significant contribution to the various firing
   response was the different levels of cellular excitability.” We
   guess that the “trivially” refers to the fact that only the
   cellular excitability has been modulated in the model. But then we
   don’t understand “A notable exception appeared for proximal
   activity: increasing proximal activity could either promote firing
   response in some cells or suppress it in some other cells whatever
   their individual excitability.” We conclude that there is still
   more than the individual excitability that determines the firing
   behavior, but this seems to contradict the previous statement. \rm

   We agree with the reviewer that the formulation "only [...] except
   [...]" might be misleading, we adapted the abstract to remove this
   source of confusion. Now it reads: \bf "but trivially, the _main_
   significant contribution to the various firing response was the
   different levels of cellular excitability [...]". \rm

#+NAME: fig:1
#+CAPTION: *Changing the ratio of the proximal/distal separation.* *(A)* Configuration of the main text, the proximal part occupies the first 87.5% of the dendritic morphology. *(B)* Now the proximal part occupies only the first 12.5% of the dendritic morphology. This importantly reduces the number of available synapses (the morphology has only one branch in the first 12.5% part). Provided this loss of synapses is compensated by an increase of activity, there is _no qualitative change between the 12.5% and the 87.5% setting for the proximal activity: non-monotonic \(\sigma_V\) variations, strong decrease in \(\tau_V / \tau_m^0 \)_. On the other hand, this change would have important qualitative consequences for the "distal activity" regime, if the last 87.5% of the morphology are "distal", many synapses are electrotonically close (what is rather irrealistic for distal synapses), they will thus stronly interact through conductance shunting and thus produce fluctuations close to the "proximal activity" setting of the main text.
[[./figs/varying_prox_dist.png]]

3) \it Main statement; missing analysis. It is suggested that the
   heterogeneity would arise from sodium channels, but the paper is
   missing a joint presentation of experimental and modeling data
   showing some fit. Neither could we find a mathematical argument why
   the different phenomenological behavior shown in the cited Fig. 5
   should arise. It would be helpful to underscore the main statement
   by a simplified mathematical formulation in the main text. For
   instance, a simplified formula that gives the output firing rate as
   a function of the input variables would be helpful, including the
   dependencies on the sodium channel density. This sodium channel
   density is claimed to make the difference between the cell types
   according to the abstract: “various levels of density of sodium
   channels and sodium inactivation will control the response of
   individual neurons to proximally targeting presynaptic activity.”
   Again, we could not find a state-of-the-art explanation why this
   statement is true. \rm

   This point is precisely the focus of our previous communication
   [[cite:Zerlaut2016]]. Unfortunately there is, to the best of our
   knowledge, no closed-form solution that could give the firing rate
   response as a response to fluctuating input in a model including
   sodium inactivation and a sodium activation curve (hence the
   phenomenological approach formulated in our previous
   communication). Nonetheless to give some intuition to the reader,
   we have added the following statement (line 316): \bf In our
   previous communication, theoretical modeling suggested that a high
   sensitivity to the speed of the fluctuations was enable by a high
   level of sodium inactivation (as only fast fluctuations allow to
   deinactivate sodium channels) and a high density of sodium channels
   (as it produces a sharp spike initiation that enables to extract
   fast fluctuating input). \rm For further details, we believe the
   reader should refer to our previous publication.

4) \it No dendritic nonlinearities. One main characteristics of L5
   pyramidal neurons is the calcium spike in the apical tree, and more
   generally the various types of dendritic spikes. Non of them is
   considered in the model, and one conjectures that the various
   response properties in the experimental data arise from these
   various dendritic nonlinearities. Describing the firing properties
   of pyramidal neurons as a function of proximal and distal activity
   should take account of these dendritic nonlinearities. Otherwise,
   in fact, the statements cannot be compelling. \rm

   We adressed this concern in details in Part A.

5) \it Data extraction. It is not clear how the 4 cells have been selected
   out of the 27 cells. The 4 cells represent 4 cell types. How are
   these cell types specified? The paper does not give an answer on
   that. \rm

   The 4 cells represent 4 different cells out of the n=30 cells
   recorded in our previous study [[cite:Zerlaut2016]]. The criteria to
   choose those 4 cells was that they should somehow represent
   different behaviors while still being representative. See in Figure
   6B in [[citetext:Zerlaut2016]] how those 4 cells cover relatively well
   the domain of measured characteristics for the firing response,
   this is also the case in the present study, where they cover
   relatively well the domain of the predicted coupling to the various
   types of activities, see Figure 6 in main text. We updated the
   manuscript to highlight this point (line 250): \bf We show on
   Figure 5 the response of four cells (representative of different
   firing response behaviors, see Figure 5 and 6B in
   [[citetext:Zerlaut2016]]) to the different types of presynaptic
   activity described in the previous section. The input-output
   relationships show qualitative and quantitative differences, [...]
   \rm

6) \it Figure 2 has two panels A & B, but the caption describes panels
   A–E. \rm

   Indeed, we apologize for this problem in the manuscript
   preparation. Figure 2 and Figure 3 have been mixed up, this is now
   corrected.

7) \it Apart from that, the fit of the numerical simulations and the
   analytical estimate in B for sigma_V is biased. No argument is
   given why this bias is there. \rm

   From line 165 to line 182, we discuss the validity of our
   analytical treatment. We explicit the two hypothesis that underlie
   our estimation and detail their expected effect. We conclude by
   discussing the bias for \(\sigma)V\) from line 180 to 182: "Some
   discrepancies between the approximation and the simulations
   appeared, in particular one can see a 1mV shift in the standard
   deviation of the fluctuation (meaning that single events are
   underestimated in the analytical treatment, so that hypothesis 2 is
   the most problematic one)." (where hypothesis 2 has been explicited
   before).

8) \it The central claim is that distal and proximal inputs differently
   affects the output firing rates. Beside the fact that this
   statement by itself is trivial [...] \rm

   We also show that "distal and proximal inputs differently
   affects the output firing rates", but this is not our main
   statement and we agree with the reviewer that this could seen as
   "trivial". Our main statement claims that those stimulations affect
   individual neurons in diverse manners (in particular for the
   proximal stimulation: increasing their response for some neurons,
   decreasing it for some other neurons). We believe that this
   diversity of responses across neurons (to the same input) is not a
   so trivial phenomena.

9) \it [...] we wondered why the closer 87.5% of the dendritic branch is
   declared as “proximal” and the furthest 12.5% as the “distal”
   part. Is the result dependent on this choice, or could one have
   also chosen the closer 12.5% as “proximal” only? \rm

   Indeed, the _arbitrary ratio_ between proximal and distal input
   might need a more detailed discussion. Following the reviewer's
   suggestion, we reproduced the analysis with the convention
   "proximal as the closest 12.5% only". We show on [[fig:1][Figure 3]] that the
   results are not affected by this change, the qualitative properties
   of the fluctuations remain identical: strong decrease of the
   fluctuations time constant and non-monotonic shape of the standard
   deviation of the fluctuations. The key property leading to the
   properties of the "proximal activity" being that they correspond to
   synapses that are close to each others and to the soma (to produce
   a strong /shunting/ effect), this phenomena is even stronger under
   this setting. The reason for choosing a relatively large fraction
   for proximal synapses is to obtain a large number of
   synapses. Indeed, along the morphology, the electrotonic distance
   raises but the surface as well (because of the branches number, see
   drawing in Fig 1 in main text) so that the number of synapses
   strongly increases. Summing up, the parameter \(f_{prox}\)=87.5% is
   important to keep the distal part far enough electrotonically, for
   the proximal part there is no much difference between
   \(f_{prox}\)=87.5% and \(f_{prox}\)=12.5% (they all contain the
   very proximal synapses), and keeping \(f_{prox}\)=87.5% avoid to
   introduce an additional rescaling of synaptic frequences because of
   the reduced number of synapses (the procedure shown in [[fig:1][Figure 3]]).

* Reviewer #2

 \it This manuscript shows a nice mixture of experiment and
 theory. The latter is based upon sensible use of the cable modelling
 framework of Rall combined with stochastic models of synaptic input
 (admittedly of a particular type but definitely fit for purpose). \rm

1) \it Unfortunately some of the work (for the firing response function)
   is buried in supplementary material - and this has not been made
   available to me (S1).\rm

   That is indeed unfortunate. For us, the permanent link at the end
   of the submission pdf (page 32/32) was functional and actually gave
   the supplementary material. Maybe the reviewer could ask the editor
   for this material if this is still not available in this
   resubmission.

2) \it It would be especially informative to hear more about the author's
   perspective on how synchrony controls rate. Nonetheless given the
   quality and novelty of the main manuscript I suggest to put this
   issue aside. \rm

   This is indeed an interesting point. In the present study, we give
   a first order description of the effect of synchrony on stationary
   firing rate (in the /weakly synchronous/ regime, close to an
   idealized /asynchronous/ regime): presynaptic synchrony between
   presynaptic events increases the variance of the fluctuations that
   in turn increases the firing rate. But there should be effect
   beyond this first order description, when we go further and further
   from the idealized asynchronous state, i.e. when the firing rate
   can not be obtained from the stationary fluctuations
   properties. Interactions between synchrony and dendritic
   non-linearities also appear important to neocortical cell
   processing (see Part A of this response). Following the reviewer's
   opinion, we would also prefer to leave this issue aside as those
   are quite complicated questions and perform detailed investigations
   in future work.

** Major concern

\it My only major concern with the paper is on the emphasis of a
theoretical framework without citing other relevant work in the
area. Essentially I am surprised not to see the mention of the
technique(s) for constructing the Green's function of an arbitrary
branched passive dendritic tree. The paper focuses on Rall equivalent
cylinders for simplicity, yet the methodology presented would seem
easily extended using the work of Abbott [[cite:Abbott1991,Cao1993]]. The
extension of the above to include a distinguished soma model can be
found in [[citetext:Coombes2007]] and Paul Bressloff has done a lot of
work on dendrites that appears relevant, some of which is summarised
in [[citetext:Bressloff1997]]. The following may also be of interest
(though not necessarily for this work) - where simple active soma
models are coupled to passive dendrites
[[cite:Bressloff1995,Svensson2009,Schwemmer2012]]. \rm

We are grateful to the reviewer for those bibliographical
suggestions. This allowed us to detail the possible extension of our
framework to more complex dendritic structures. The updated discussion
now reads (line 335): \bf To investigate those effects within the
framework proposed in our study, one could benefit from the large body
of theoretical work on the derivation of Green’s function for
arbitrary branched passive dendritic trees
[[cite:Abbott1991,Cao1993,Bressloff1997,Coombes2007]]. \rm

Nonetheless, for clarity for the reader (not to be pointed toward more
complex methods than just Rall's method), we keep only the reference
to Rall's work in the Methods and Results section. Then the author can
be more focused on the originality of our approach: the use of
shotnoise theory results in a dendritic integration context.

In addition, we have given some perspective about the approach used
here to obtain the firing rate response in presence of a dendritic
structure. In the discussion, one can now find the sentence (line
361):

\bf This phenomenological two-steps procedure thus offers a flexible
complementary approach to the analytical approaches tackling the
problem of the spiking behavior in presence of an extended dendritic
structure [[cite:Svensson2009,Schwemmer2012,Saparov2015]]. \rm

** Minor points

1) \it Spurious "s" in the sec "A simplified morphological model" - sWe arbitrarily ... \rm

Thanks. corrected.

2) \it Where is excitability < V_thre^eff >_D defined? \rm

Thanks. This was indeed only briefly mentionned in a main text as a
reference to our previous study, now in the Methods section, we added:

\bf Importantly, our study introduced four quantities to describe the
mean properties of a single neuron response in the fluctuation-driven
regime: (i) a measure of the excitability given by the mean
phenomenological threshold \( \langle V_{thre}^eff \rangle_\mathcal{D}
\) of our firing rate function, (ii) a sensitivity to the mean of the
fluctuations \( \langle \frac{\partial \nu}{\partial \mu_V}
\rangle_\mathcal{D} \), i.e. quantifying how much a mean
depolarization is translated into a change in firing rate, (iii) a
sensitivity to the amplitude of the fluctuations \( \langle
\frac{\partial \nu}{\partial \sigma_V} \rangle_\mathcal{D} \),
quantifying how much an increase of the standard deviation of the
fluctuations is translated into a change in firing rate, and (iv) a
sensitivity to the speed of the fluctuations \( \langle \frac{\partial
\nu}{\partial \tau_V^N} \rangle_\mathcal{D} \), quantifying how much a
change in the autocorrelation time of the fluctuations is translated
into a change in firing rate. \rm

* Reviewer #3

\it The key claim of this work is that recently identified L5 PCs firing
rate properties indicate how these cells couple to presynaptic
activity patterns (proximal, distal,...). The claim is based on i)
recently identified firing response functions (previous work of the
authors), ii) a theoretical model for passive dendritic integration
that is an extension of the single compartment model of
[[citetext:Kuhn2004]] to a passive tree with cable theory, and iii) an
identification of the passive tree of L5 PCs by fitting the input
impedance at the soma to experimental data.

The topic is of importance and the theoretical part is novel, very
appealing in itself and rigorous, but inadequate to support the key
claim of the paper (see paper title).

This is because ... \rm

** CONS

 - *A*: \it (part ii) For L5 PCs it is highly questionable if dendritic
   integration can be understood without (or separated from) active
   dendritic properties given the known experimental facts and huge
   literature, e.g. about the compartmentalization of dendritic
   function, dendritic spikes, and changed response properties due to
   ion channel blockers.

   In my opinion, the model for the L5 PCs based on a passive tree
   will almost surely fail to describe the real dendritic integration
   properties of the cells in the fluctuation driven regime. In any
   case, see the first point in additional work below. \rm

   We adressed in details the question of dendritic nonlinearities in
   Part A. The question of the compartementalization of dendritic
   function is indeed something not adressed in our study, we would
   prefer to leave this aspect for future investigations.

- *B*: \it (part iii) Even the description of the passive dendritic
  integration properties of the cell seems to be questionable (maybe
  because of the constraints to symmetrical trees). That this doesn't
  work out properly seems to be reflected in the morphologies obtained
  in the paper: Is the required variation of the tree length from
  about 400 um to 900 um (current Fig 3C and Fig. S1) for PCs
  plausible? Please provide references. L5 PCs with their long apical
  dendrites etc. are the worst case scenario for the symmetry
  constraints for the tree. \rm

  We totally agree with the reviewer on this point. We removed the
  point stating that our simplified morphology could somehow resemble
  those of layer V pyramidal cells. More precisely, we removed the
  following statement: \bf Also the obtained morphology seems
  realistic (B=5 generation of branches, a total tree length of =550m
  and a root branch diameter of the tree =2.25m) given what can be
  observed after histological work [[cite:Zerlaut2016]] \rm. The emphasis
  should remain on the electrical properties.

  Importantly, the results of our study do not rely on having a
  precise description of the morphology, we only use this to
  discriminate between a proximal and a distal area, and Rall's
  symmetric branching model is sufficient to reproduce this feature.
  So in the discussion, we aknowledge the important limitations of our
  morphological model and stress the above point (what is the
  important ingredient in our study) with:

 \bf [...] even if our description captures the various electrotonic
  distances associated to various synaptic locations (the crucial
  ingredient here to discriminate between proximal and distal inputs),
  the morphological model appears as a _very poor description of layer
  V pyramidal cell_. Deviations from the symmetric branching
  hypothesis and Rall's branching rule will have a significant impact
  on dendritic integration in the fluctuation-driven regime. To
  investigate those effects [...] \rm

** PROS

\it The theoretical work is still significant for other cell types that
work in a more linear regime, e.g. L2/3 PCs, and have morphologies
that match the theoretical constraints.. \rm

** Additional work required:

- \it The authors should provide evidence that active dendritic
  integration doesn't change the key claim, or, apply the model to a
  more appropriate cell type. \rm

  see Part A of this response

- \it Because the morphology of the tree of the model is very
  constraint and only a very rough estimate its necessary to assess
  the influence of this approximation on dendritic integration: A
  systematic analysis of the impact of the morphology of the tree (7
  variables) on the response properties of interest at the soma is
  missing (This allows also a generalization of the results to other
  cell types. Now only Fig S2 gives a hint about variability). \rm

  *TO BE DONE*

- \it There's no reference to any other model than [[citetext:Kuhn2004]],
  on which the approach is based. Relate the model to recent work on
  the effects of passive dendritic tree properties of neuron on firing
  dynamics (e.g.  [[citetext:Saparov2015]] or other model reduction
  techniques (a huge field, e.g. [[citetext:Hedrick2013]]). \rm

  The updated discussion now reads (line 335): \bf To investigate
  those effects within the framework proposed in our study, one could
  benefit from the large body of theoretical work on the derivation of
  Green’s function for arbitrary branched passive dendritic trees
  [[cite:Abbott1991,Cao1993,Bressloff1997,Coombes2007]]. \rm In addition,
  we have given some perspective about the approach used here to
  obtain the firing rate response in presence of a dendritic
  structure. In the discussion, one can now find the sentence (line
  361): \bf This phenomenological two-steps procedure thus offers a
  flexible complementary approach to the analytical approaches
  tackling the problem of the spiking behavior in presence of an
  extended dendritic structure
  [[cite:Svensson2009,Schwemmer2012,Saparov2015]]. \rm

\it The paper is well organized, easy to read, simple to follow and
clearly written.

However, there are too many typos and word repetitions to list them
all ("sWe" p.3, "show deviations from the deviations" p.4, " vary the
size of the size" p.4, "does not goes", p.6, "the firing increase
show", p.7 ...) and the Figure numbers were incorrect (Fig.22b, p.4)
or mixed up (Fig. 2 and 3). \rm

We apologize for the problems in the manuscript preparation (the
figures mixed up) and we thank the reviewer for pointing out the
typos... From now on, even if the question refer to the (wrong)
previous notation ex-Figure 3 will be renamed to Figure 2 as in the
re-submission.

** Minor:

- \it Fig. 3: The color code seems to be inconsistent: Panel A shows cells
  with Rm above 600 MOhm, but the bottom panel in Fig 3C shows that
  such cells are not present. \rm

  This is because those data come from two separate datasets, we
  highlighted this with the mention of "firing rate dataset" at the
  bottom of panel 2C. We have a special dataset to investigate the
  input impedance properties (on the exact same experimental model)
  where 2 cells were more resistive than those present in the other
  dataset. We use this dataset to derive a "sizing rule" from the
  input resistance and then be able to use the "firing rate dataset"
  where we only have the input resistance (and not the detailed input
  impedance spectrum). Note that, as expected from the same
  exerimental system, the datasets were statistically pretty similar:
  355.9 \(\pm\) 184.1 M\(\Omega\) for the "firing rate dataset", see
  [[citetext:Zerlaut2016]] and 387.3 \(\pm\) 197.2 M\(\Omega\) for the
  "input impedance dataset", see Methods.

  We have updated figure 2C with the mention of "n=30cells" for the
  "firing rate dataset", then it highlights the difference with the
  dataset of input impedance labeled "n=13cells" and makes it more
  clear to the reader.

- \it Fig 3C: The lines in the three top panels could be drawn with the
  color code (color changes within a line) as used in panel A and B. \rm

  Very good idea, thanks ! See new Figure 2.

- \it Fig 3AB: Please comment on the inconsistency between data and model
  for > 100 Hz. There seems to be a flip in vertical order. \rm

  *TO BE DONE*

\it Section: "An analytical approximation ...":

- The claim "the analytical estimate could therefore be seen as an
  accurate estimate, modulo a shift in the synchrony (~0.2 increase in
  synchrony corrects for the ~1 mV shift in sigma)" should
  verified: 1) Results for synchrony = 0.2 should be shown in the
  right most column of panels in Fig 2B, and 2) results for synchrony
  = 0.2 should be shown for the other columns of panels in Fig 2B,
  where the other parameters are varied. \rm

  *TO BE DONE*

Section: "Biophysical origin ..."

- \it The characteristics of the firing response function are not properly
  introduced. Please add this. \rm

  This is indeed a very important point, thanks, we updated the
  methods section with:

  \bf Importantly, our study introduced four quantities to describe
  the mean properties of a single neuron response in the
  fluctuation-driven regime: (i) a measure of the excitability given
  by the mean phenomenological threshold \( \langle V_{thre}^eff
  \rangle_\mathcal{D} \) of our firing rate function, (ii) a
  sensitivity to the mean of the fluctuations \( \langle
  \frac{\partial \nu}{\partial \mu_V} \rangle_\mathcal{D} \),
  i.e. quantifying how much a mean depolarization is translated into a
  change in firing rate, (iii) a sensitivity to the amplitude of the
  fluctuations \( \langle \frac{\partial \nu}{\partial \sigma_V}
  \rangle_\mathcal{D} \), quantifying how much an increase of the
  standard deviation of the fluctuations is translated into a change
  in firing rate, and (iv) a sensitivity to the speed of the
  fluctuations \( \langle \frac{\partial \nu}{\partial \tau_V^N}
  \rangle_\mathcal{D} \), quantifying how much a change in the
  autocorrelation time of the fluctuations is translated into a change
  in firing rate. \rm

- \it It can be hardly claimed that Fig. 6A is approximately a log normal
  distribution. \rm

  We agree. What we actually mean is that the data appears to be
  distributed logarithmically, we do not wish to put a special
  emphasis on the log-normal trend. The statement "approximatelly
  log-normal" was a concise way to express this trend. If this is a
  problem, we could change it to "the data appears to be distributed
  logarithmically and not linearly" but we are not sure that this
  formulation would render better the idea.

\it Methods section: \rm

- \it Not all variables are properly defined, e.g. Qe, Qi, lp. \rm

  Indeed, thanks. We updated the Methods section (mostly the "Model
  equations" section).

- \it "it constitutes an extension of the approach proposed in Kuhn et
  al.": Please, state what extension. \rm

  Done, it now reads: \bf it extends the approach proposed in Kuhn et
  al. to the extended dendritic structure following Rall’s branching
  rule \rm

- \it "Fitting passive properties": Are the solutions unique or are there
  many different trees with similar fit performance that result in
  different couplings to presynaptic activity? \rm

  *TO BE DONE*

* References
\small

#+BEGIN_LATEX
\begin{filecontents}{biblio.bib}

@article{Hausser2005,
author = {London, Michael and H{\"{a}}usser, Michael},
doi = {10.1146/annurev.neuro.28.061604.135703},
issn = {0147-006X (Print)$\backslash$r0147-006X (Linking)},
journal = {Annual review of neuroscience},
pages = {503--532},
title = {{Dendritic computation.}},
type = {Journal Article},
volume = {28},
year = {2005}
}

@article{Stuart2015,
author = {Stuart, Greg J and Spruston, Nelson},
doi = {10.1038/nn.4157},
file = {:Users/yzerlaut/Documents/Mendeley Desktop/Stuart, Spruston - 2015 - Dendritic integration 60 years of progress.pdf:pdf},
issn = {1097-6256},
journal = {Nature Neuroscience},
mendeley-groups = {Neuroscience},
number = {12},
pages = {1713--1721},
pmid = {26605882},
title = {{Dendritic integration: 60 years of progress}},
url = {http://www.nature.com/doifinder/10.1038/nn.4157},
volume = {18},
year = {2015}
}

@article{Larkum2009,
author = {Larkum, M E and Nevian, T and Sandler, M and Polsky, A and Schiller, J},
doi = {10.1126/science.1171958},
issn = {1095-9203 (Electronic) 0036-8075 (Linking)},
journal = {Science},
number = {5941},
pages = {756--760},
title = {{Synaptic integration in tuft dendrites of layer 5 pyramidal neurons: a new unifying principle}},
type = {Journal Article},
url = {http://www.ncbi.nlm.nih.gov/pubmed/19661433 http://science.sciencemag.org/content/sci/325/5941/756.full.pdf},
volume = {325},
year = {2009}
}

@article{Hedrick2013,
author = {Hedrick, K R and Cox, S J},
doi = {10.1007/s10827-012-0403-y},
issn = {1573-6873 (Electronic) 0929-5313 (Linking)},
journal = {J Comput Neurosci},
number = {1},
pages = {1--26},
title = {{Structure-preserving model reduction of passive and quasi-active neurons}},
type = {Journal Article},
url = {https://www.ncbi.nlm.nih.gov/pubmed/22714391},
volume = {34},
year = {2013}
}

@article{Saparov2015,
  title={Effects of passive dendritic tree properties on the firing dynamics of a leaky-integrate-and-fire neuron},
  author={Saparov, Abulhair and Schwemmer, Michael A},
  journal={Mathematical biosciences},
  volume={269},
  pages={61--75},
  year={2015},
  publisher={Elsevier}
}

@article{Schwemmer2012,
  title={Bistability in a leaky integrate-and-fire neuron with a passive dendrite},
  author={Schwemmer, Michael A and Lewis, Timothy J},
  journal={SIAM Journal on Applied Dynamical Systems},
  volume={11},
  number={1},
  pages={507--539},
  year={2012},
  publisher={SIAM}
}

@article{Bressloff1995,
  title={Dynamics of a compartmental model integrate-and-fire neuron with somatic potential reset},
  author={Bressloff, Paul C},
  journal={Physica D: Nonlinear Phenomena},
  volume={80},
  number={4},
  pages={399--412},
  year={1995},
  publisher={Elsevier}
}

@article{Svensson2009,
author = {Svensson, C.-M. and Coombes, S},
doi = {10.1142/S0218127409024347},
journal = {International Journal of Bifurcation and Chaos},
pages = {2597--2607},
title = {{Mode Locking in a Spatially Extended Neuron Model: Active Soma and Compartmental Tree}},
type = {Journal Article},
volume = {19},
year = {2009}
}

@article{Bressloff1997,
author = {Bressloff, P. C. and Coombes, S.},
doi = {10.1142/S0217979297001209},
file = {:Users/yzerlaut/Documents/Mendeley Desktop/Bressloff, Coombes - 1997 - Physics of the Extended Neuron.pdf:pdf},
issn = {0217-9792},
journal = {International Journal of Modern Physics B},
mendeley-groups = {Neuroscience},
month = {aug},
number = {20},
pages = {2343--2392},
title = {{Physics of the Extended Neuron}},
url = {http://www.worldscientific.com/doi/abs/10.1142/S0217979297001209},
volume = {11},
year = {1997}
}

@article{Coombes2007,
author = {Coombes, S and Timofeeva, Y and Svensson, C-M and Lord, G J and Josi{\'{c}}, K and Cox, S J and Colbert, C M},
doi = {10.1007/s00422-007-0161-5},
file = {:Users/yzerlaut/Documents/Mendeley Desktop/Coombes et al. - 2007 - Branching dendrites with resonant membrane a sum-over-trips approach.pdf:pdf},
issn = {0340-1200},
journal = {Biological cybernetics},
keywords = {Action Potentials,Action Potentials: physiology,Algorithms,Animals,Cell Membrane,Cell Membrane: physiology,Cell Shape,Cell Shape: physiology,Computer Simulation,Dendrites,Dendrites: physiology,Hippocampus,Hippocampus: physiology,Membrane Potentials,Membrane Potentials: physiology,Models, Neurological,Rats,Synaptic Transmission,Synaptic Transmission: physiology},
mendeley-groups = {Neuroscience},
month = {aug},
number = {2},
pages = {137--49},
pmid = {17534649},
title = {{Branching dendrites with resonant membrane: a "sum-over-trips" approach.}},
url = {http://www.ncbi.nlm.nih.gov/pubmed/17534649},
volume = {97},
year = {2007}
}

@article{Cao1993,
author = {Cao, B J and Abbott, L F},
doi = {10.1016/S0006-3495(93)81370-5},
issn = {0006-3495 (Print) 0006-3495 (Linking)},
journal = {Biophys J},
number = {2},
pages = {303--313},
title = {{A new computational method for cable theory problems}},
type = {Journal Article},
url = {https://www.ncbi.nlm.nih.gov/pubmed/8457662},
volume = {64},
year = {1993}
}

@article{Abbott1991,
author = {Abbott, L F and Farhi, E and Gutmann, S},
issn = {0340-1200 (Print) 0340-1200 (Linking)},
journal = {Biol Cybern},
number = {1},
pages = {49--60},
title = {{The path integral for dendritic trees}},
type = {Journal Article},
url = {https://www.ncbi.nlm.nih.gov/pubmed/1768712},
volume = {66},
year = {1991}
}

@article{Zerlaut2016,
author = {Zerlaut, Y and Telenczuk, B and Deleuze, C and Bal, T and Ouanounou, G and Destexhe, A},
doi = {10.1113/JP272317},
issn = {1469-7793 (Electronic) 0022-3751 (Linking)},
journal = {J Physiol},
number = {13},
pages = {3791--3808},
title = {{Heterogeneous firing rate response of mouse layer V pyramidal neurons in the fluctuation-driven regime}},
type = {Journal Article},
url = {http://www.ncbi.nlm.nih.gov/pubmed/27146816},
volume = {594},
year = {2016}
}

@article{Kuhn2004,
abstract = {During sensory stimulation, visual cortical neurons undergo massive synaptic bombardment. This increases their input conductance, and action potentials mainly result from membrane potential fluctuations. To understand the response properties of neurons operating in this regime, we studied a model neuron with synaptic inputs represented by transient membrane conductance changes. We show that with a simultaneous increase of excitation and inhibition, the firing rate first increases, reaches a maximum, and then decreases at higher input rates. Comodulation of excitation and inhibition, therefore, does not provide a straightforward way of controlling the neuronal firing rate, in contrast to coding mechanisms postulated previously. The synaptically induced conductance increase plays a key role in this effect: it decreases firing rate by shunting membrane potential fluctuations, and increases it by reducing the membrane time constant, allowing for faster membrane potential transients. These findings do not depend on details of the model and, hence, are relevant to cells of other cortical areas as well.},
author = {Kuhn, Alexandre and Aertsen, Ad and Rotter, Stefan},
doi = {10.1523/JNEUROSCI.3349-03.2004},
file = {:Users/yzerlaut/Documents/Mendeley Desktop//Kuhn, Aertsen, Rotter - 2004 - Neuronal integration of synaptic input in the fluctuation-driven regime.pdf:pdf},
issn = {1529-2401},
journal = {The Journal of neuroscience : the official journal of the Society for Neuroscience},
keywords = {Action Potentials,Action Potentials: physiology,Animals,Computer Simulation,Excitatory Postsynaptic Potentials,Excitatory Postsynaptic Potentials: physiology,Humans,Membrane Potentials,Membrane Potentials: physiology,Models,Neural Inhibition,Neural Inhibition: physiology,Neurological,Neurons,Neurons: physiology,Synapses,Synapses: physiology,Synaptic Transmission,Synaptic Transmission: physiology,Visual Cortex,Visual Cortex: physiology},
mendeley-groups = {Neuroscience},
month = {mar},
number = {10},
pages = {2345--56},
pmid = {15014109},
publisher = {Soc Neuroscience},
title = {{Neuronal integration of synaptic input in the fluctuation-driven regime.}},
url = {http://www.ncbi.nlm.nih.gov/pubmed/15014109},
volume = {24},
year = {2004}
}
\end{filecontents}
#+END_LATEX

\bibliography{biblio}

* Options :noexport:
# #+OPTIONS: toc:nil num:1
#+OPTIONS: num:1
#+LaTeX_CLASS_OPTIONS: [8pt, a4paper, colorlinks]
#+LaTeX_HEADER: \hypersetup{allcolors = blue} % to have all the hyperlinks in 1 color
#+LaTeX_HEADER: \usepackage{amssymb,mathenv,array}
#+LaTeX_HEADER: \usepackage{microtype} % Slightly tweak font spacing for aesthetics
#+LaTeX_HEADER: \usepackage{geometry}
#+LaTeX_HEADER: \geometry{a4paper,total={210mm,297mm}, left=25mm, right=25mm, top=20mm, bottom=20mm}
#+LaTeX_HEADER: \usepackage{filecontents}
#+LaTeX_HEADER: \renewcommand{\refname}{\vspace{-.8cm}}
#+LaTeX_HEADER: \usepackage{biology_citations}
#+LaTeX_HEADER: \bibliographystyle{biology_citations}
#+LaTeX_HEADER: \renewcommand{\thesection}{\Alph{section}}